### 2012-01-09 Paper_honeypot.tex
setup gedeeltelijk bijgewerkt

\documentclass[11pt]{article}
\usepackage{a4wide, graphicx, fancyhdr, wrapfig, tabularx, amsmath, amssymb, hyperref, color, verbatim, nameref}
\usepackage[english]{babel}
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour, linkcolor=linkcolour}

%----------------------- Macros and Definitions --------------------------

\setlength\headheight{20pt}\usepackage{}
\addtolength\topmargin{-10pt}
%\addtolength\footskip{20pt}

\fancypagestyle{plain}{%
\fancyhf{}
\fancyfoot[RO,LE]{\sffamily\bfseries\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot[RO,LE]{\sffamily\bfseries\thepage}
\fancyhead[RO,LE]{\textsc{}}
\fancyhead[LO,RE]{\emph{}}
\renewcommand{\headrulewidth}{1pt}
\renewcommand{\footrulewidth}{0pt}
\newcommand{\tab}{\hspace*{2em}}

\newcommand{\tocheck}[1]{{\bf !?: #1 :!?}}
%\newcommand{\OBA}{Online behavioural advertising }

\frenchspacing

%-------------------------------- Title ----------------------------------

\title{\textbf{Honeypot Assignment: \\ \emph{Group Contagiarium}}}
\author{Aram Verstegen, \tocheck{studentnummer Utwente} \\
	 Mark Vijfvinkel, 1275194 \\
	 Bart Lutgens, \tocheck{studentnummer Utwente} \\
	Joost Kremers, \tocheck{studentnummer Utwente}}

\date{\today}

\begin{document}
\maketitle

\section{Introduction}
\tocheck{Kijken of the secties nog kloppen!}
In this paper we will try to present an overview of the data we have collected from the honeypot and explain the different attacks that were waged against it. 
In the second section we will describe the changes we have made to the setup during the "Waiting for hackers" phase. 
The third section will present an overview of various statistics we have collected from the logfiles, e.g. number of IP-addresses, countries and attack types. 
In the fourth section we will discuss which attacks took place against the different services the honeypot was running e.g. SSH and MySQL.
In order to discover the sophisticationof the attacks, we will try to present possible connections between an attack on one service and an attack on a different service, in the fifth section.
The sixth section will conclude our findings. 
Finally we will discuss this project and provide some improvements.

\section{Setup}
\label{Setup}

On the 1st of November we delivered our initial setup document, however during the "Waiting for hackers" phase we have made adjustments to the setup in order to attract more activity to the honeypot. In this section we will present the initial setup and in addittion to that, discuss the changes that were made.

%%%%MOET NOG WORDEN HERZIEN%%%%%

\subsection{Monitoring Machine}

In this subsection we describe the setup of the monitoring machine.

\subsubsection{Syslog-ng}
By default, \verb|syslog| uses UDP to transmit its log messages to a remote server.
Relying on UDP is not desirable, because everything is sent in plaintext and delivery of UDP messages is not guaranteed.
To allow for confidential and reliable transmission of the syslog messages from the one machine to another, one needs to use TCP connections in syslog, and provide some way of encapsulating the socket used in a cryptographic layer.

We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with SSH and syslog-ng''. \cite{remote_logging}\tocheck{Toevoegen refs}
The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.

We have configured a locally installed \verb|syslog-ng| package for our user on the monitoring machine and have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
%We are aware there is still the possibility of other users on the monitoring machine poisoning our logs, but we feel we can rule out foul play by our peers in this assignment.

To allow for secure SSH access, a public keypair was generated on the util machine and the associated public key was added to the honeypot's \verb|authorized_keys| file.

We have put the command to create the SSH tunnel in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason. 
A severance of the connection would lead us to assume the worst initially.
If re-establishing this connection fails after any attackers have gained access (which should never happen if we monitor adequately), we would have assumed that the machine was compromised beyond our control and would have requested a shutdown of the VM, if no network downtime was otherwise announced.

\subsubsection{Monitoring}
After collecting our syslog information, we wanted to be somehow alerted of interesting information automatically.
We decided to use an IRC bot that would leave a message in an IRC channel (\#contagiarium) everytime an entry was made to the syslog file. First this lead to an overloaded channel and several false positives. To solve these problems filters were created to only post relevant information. For example, from the kippo logs only the actual login attempts would be shown with the tried password.

\subsection{Honeypot}

In this section we describe the setup of the honeypot.

\subsubsection{Rsyslog}
On the honeypot, we modified the default \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain to the syslog instance on our monitoring machine.
The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking an interest.
%%Bespreken Aram
We plan to move this small configuration over to one of the default configuration files to avoid raising suspicion outright.
Note that the chosen port for remote syslog is very similar to the default syslog port 514, and this might give us away.
We plan to change it to a more nondescript port number.

\subsubsection{SSHFS mount}
Not all services support logging to syslog, and for some it is infeasible.
We created a read-only \verb|sshfs| mount from our monitoring machine that we will use to write our logfiles for various services to.
We used interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.

\subsubsection{scanlogd}
To detect port scans we have installed the service \verb|scanlogd|.
Per detected scan an entry is written to syslog.

\subsubsection{iptables}
To log all incoming and outgoing connection attempts, we have added the following rules to the iptables configuration. The attempts were also logged to syslog.

\begin{verbatim}
iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "
iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "
\end{verbatim}

%Bespreken Aram
%We are also going to filter outgoing SMTP, SIP and NetBIOS traffic.

\subsubsection{Kippo}
To capture attackers that try to come in by brute-forcing SSH logins, we are used Kippo. 
Kippo is a sandbox SSH service, used to log bruteforce attacks and the shell interaction performed by the attacker after gaining access (with a trivial password). 
The latest development version was installed to run as a separate user.

To make this as realistic as possible we added a redirect in the iptables configuration from the real SSH port 22 to port 2222 that Kippo uses by default.
The reason for forwarding in iptables was to not give Kippo root rights, which would be very dangerous.
%Bespreken Aram
Unfortunately the forward broke down around the time the firewall was being reconfigured, for reasons yet unclear . 
If it can't be resolved, we can circumvent this issue by running Kippo through a copy of the Python interpreter that has the capability to bind privileged ports.

To be able to still log on to the honeypot, we configured our real SSH service to run on port 9022.

Furthermore, we patched Kippo to log messages to syslog.

During the second phase of the project, we noticed that of the succesfull bruteforce attempts against Kippo, none actually tried to continue and explore or use the honeypot. Running a default portscan with Metasploit revealed that it can detect if Kippo is being used. We thought this explained why nobody tried to log in. To solve this problem, we adjusted the Kippo configuration so that Metasploit could recognize Kippo. Unfortunately did this not lead to actual exploration or use of the honeypot.

\subsubsection{DenyHosts}
If someone found out that our real SSH port was changed to 9022, they might try to break in through the real SSH service.
To make sure no bruteforce attacks can be done on this port, DenyHosts was installed.
This service blocks IP-addresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
The hosts listed there are blacklisted on the machine, but we can edit them if needed.
%bespreken Aram
Denyhosts can log to syslog by editing its configuration file.

\subsection{Apache \& Nginx}
To attract web traffic, we installed the Apache web server, the Apache PHP runtime and the \verb|mod-security| module which allows us to spoof its version information, to make it appear old and vulnerable.
The Nginx webserver is installed as a reverse proxy in front of Apache, which runs on local port 8080.
This provides 2 layers of logging (to files), and Nginx's superior transparent caching mechanism allows us to avoid exhausting our resources with scripting languages, should we have ever faced an application-layer DoS attack (Nginx appears quite resilient against those).

Nginx has been compiled from source and was patched to send the spoofed Apache version header as well.
We will configure both services to log to our \verb|sshfs| mount.

\subsubsection{Google Hack Honeypot}
Google Hack Honeypot (GHH) is designed to be used to attract search engine attackers. 
These attackers will use search queries to find sensitive data, for example password files, that due to some misconfiguration ended up being indexed by a search engine like Google.
GHH allows logging of these requests to a MySQL database or Comma Seperated Value (CSV) file.

Since GHH offers multiple search targets, we picked one to start with, namely GGH version 1.2 passlist.txt.
We had no particular reason for this, only that it was easy to set up. 
%We plan to install more variants of the GHH to attract more attackers, but before this could happen the VM's were shut down.

%To set up GHH we have first installed Apache and PHP5.
We configured GHH to log to in CSV format, since it requires minimal setup and is easily parsed.
%We can later reconfigure to use MySQL, if the results are not satisfactory.
Then, we patched the GHH code to allow logging to syslog, to maintain a single, consistent point of log collection.

We hoped to detect and capture a bruteforce attack or scan on generic filenames, for example \verb|passwd.txt|, because this would probably have warned us that an attacker will try to use a dictionary attack.
Unfortunately GGH broke down after the Virtual Machines were shut down, we were unable to get it back running and therefore decided not invest more time and effort into it. We also believe it would not have brought us more interesting data.

\subsubsection{Website}

\subsection{Wordpress}

To capture more attackers during the HI phase, we plan to use an outdated version of WordPress. 
Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system. 
Exploits are found on a frequent basis for Wordpress and patches are published on a regular basis. 
We plan to use an older version of Wordpress, for example a version below version 3.0.4, to be sure that we use a version for which public exploits exist.

\subsection{Snort}
Snort is an open source intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
Snort was designed to be used as a network IDS, however we will be running it on our host machine, effectively making it a host-based IDS.

Unfortunately the VM network went down before we were able to complete our Snort configuration.
We intend to set up Snort using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
Included are rules on detecting exploit attempts on various common services like HTTP, FTP, SMTP, and many more.

Considering we are in a VM we might be limited in RAM, which Snort seems to require a lot of.
The configured rules will have to be minimal to avoid exhausting resources.

We will configure Snort to send alerts to syslog, and to log its packet captures to the read-only \verb|sshfs| mounted from our monitoring machine.
We will install a script like \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.

\subsection{Auditing}
For the HI phase, we would like to have a way to audit commandline interaction. 
Various options exist, such as the user-space \verb|auditd| daemon, but to make auditing harder to detect this we feel it would be best to look for a kernel-space solution.
An auditing module developed by the Honeynet project called Sebek can provide this, but is unmaintained and outdated.
The grsec kernel patches also provide auditing from kernel space, but we were advised against replacing the stock kernel.

We are still investigating the best way to achieve useful auditing.
A self-written kernel module that hooks system calls, albeit limited, might suffice to get some much needed auditing information from the honeypot.

\section{Statistics}
\label{Statistics}
In this section we will discuss the statistical data that have been gathered from our logfiles.

\section{Attacks on the different services}
\label{Attacks}
During phase 2 we have monitored the honeypot closely and collected data about different attacks against the honeypot. In this section we will discuss the attacks against the services the honeypot was running.

\section{Connection between the attacks}
\label{Connection}
In this section we will correlate the data about attacks against one service against the data of attacks of a different service. We will try to look for a pattern or similarities in this data and determine if the attack was more sophisticated.

\section{Conclusion}
\label{Conclusion}
From Section \ref{Connection} we can conclude that...

\section{Discussion}
\label{Discussion}
In this section we would like to discuss some issues we encountered during the honeypotproject.

\begin{itemize}
\item Low/High interaction phase
\item GHH
\item etc
\end{itemize}

%%% REFS %%%

\bibliographystyle{plain} % amsalpha
\bibliography{assignment2}

\end{document}