### 2012-01-10 Paper_honeypot.tex

\documentclass[11pt]{article}
\usepackage{a4wide, graphicx, fancyhdr, wrapfig, tabularx, amsmath, amssymb, hyperref, color, verbatim, nameref}
\usepackage[english]{babel}
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour, linkcolor=linkcolour}

%----------------------- Macros and Definitions --------------------------

\setlength\headheight{20pt}\usepackage{}
\addtolength\topmargin{-10pt}
%\addtolength\footskip{20pt}

\fancypagestyle{plain}{%
\fancyhf{}
\fancyfoot[RO,LE]{\sffamily\bfseries\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot[RO,LE]{\sffamily\bfseries\thepage}
\fancyhead[RO,LE]{\textsc{}}
\fancyhead[LO,RE]{\emph{}}
\renewcommand{\headrulewidth}{1pt}
\renewcommand{\footrulewidth}{0pt}
\newcommand{\tab}{\hspace*{2em}}

\newcommand{\tocheck}[1]{{\bf !?: #1 :!?}}
%\newcommand{\OBA}{Online behavioural advertising }

\frenchspacing

\usepackage{booktabs}
\usepackage[parfill]{parskip}
%-------------------------------- Title ----------------------------------

\title{\textbf{Honeypot Assignment: \\ \emph{Group Contagiarium}}}
\author{Aram Verstegen, s4092368 \\
	 Mark Vijfvinkel, 1275194 \\
	 Bart Lutgens, 12752408 \\
	Joost Kremers, 1275259}

\date{\today}

\begin{document}
\maketitle

\section{Introduction}
\tocheck{Kijken of the secties nog kloppen!}
In this paper we will try to present an overview of the data we have collected from the honeypot and explain the different attacks that were waged against it. 
In the second section we will describe the changes we have made to the setup during the "Waiting for hackers" phase. 
The third section will present an overview of various statistics we have collected from the logfiles, e.g. number of IP-addresses, countries and attack types. 
In the fourth section we will discuss which attacks took place against the different services the honeypot was running e.g. SSH and MySQL.
In order to discover the sophistication of the attacks, we will try to present possible connections between an attack on one service and an attack on a different service, in the fifth section.
The sixth section will conclude our findings. 
Finally we will discuss this project and provide some improvements.

\section{Setup}
\label{Setup}

On the 1st of November we delivered our initial setup document, however during the "Waiting for hackers" phase we have made adjustments to the setup in order to attract more activity to the honeypot. In this section we will present the initial setup and in addition to that, discuss the changes that were made.

%%%%MOET NOG WORDEN HERZIEN%%%%%

\subsection{Rsyslog}
At first the rsyslog configuration was stored in its own file in the syslog.d directory, which might have been seen by an attacker.
Therefore we have moved it to the default syslog configuration file.

\subsubsection{Kippo}
During the second phase of the project, we noticed that of the succesful dictionary-based bruteforce attempts against Kippo, none actually tried to continue and explore or use the honeypot.
Running an \verb|nmap| portscan from the Metasploit framework (which adds to the information provided by \verb|nmap|) revealed that it can be detected if Kippo (or rather Python's twisted framework) is being used.
We learned that twisted can be identified because it sends the supported ciphers during the key exchange phase before it has been prompted to do so.
Still, this doesn't explain why people should refrain from logging in after a successful dictionary attack, as a sophisticated attacker would not pursue a dictionary attack if he were aware of the service acting as a honeypot.

\subsubsection{Google hack honeypot}
We hoped to detect and capture a brute force attack or scan on generic filenames, for example \verb|passwd.txt|, because this would probably have warned us that an attacker will try to use a dictionary attack.
In retrospect, given that our honeypot does not provide an actual service, it is easy to identify dictionary-based scans in our logs.
Unfortunately our configured GGH broke after an upgrade, and we decided not to spend more time on configuring it.
Through other teams we became aware of the HIHAT software package, which might be a more useful alternative.
%TODO references?

\subsubsection{Website}
To attract more activity we decided to create a fake website that would resemble a website of the University of Twente. To make it more appealing in ways of computing power we assumed the fictional identity of ``The department of Partical Physics and Astronomy of the University of Twente'' as a cover.
On the website we also put some information about a fictive person (Jenny Abbott) would manage the servers and website. So we created a user account with the same login and password, namely Jenny. This way a more sophisticated attack would be possible. An attacker would look at the website, determine if it is interesting to attack and then connect to Kippo to log in as Jenny.

\subsubsection{Wordpress}
In order to make the website more vulnerable and appealing to hackers, we installed an out of date version of Wordpress. Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system. Exploits are found on a frequent basis for Wordpress and therefore patches are published frequently. In the first week of December we installed version 2.8.4 of Wordpress on the honeypot.

\subsubsection{Snort and Suricata}
Unfortunately we noticed that snort required a lot of resources and was slowing down the honeypot. We replaced snort with suricata, which is compatible with the snort ruleset format.
This also allowed us to keep up with the snort community rules, which were refreshed daily using \verb|oinkmaster|.
Furthermore, Suricata allows use of more advanced packet capture interfaces than \verb|libpcap|.
We've configured suricata to sniff packets using the PF\_RING capture interface, a kernel-mode ring buffer of packets exposed via \verb|procfs|.
This provided great results and allowed us to use many rules on a limited amount of memory.

\subsubsection{Dionaea}
Dionaea is a low interaction honeypot daemon that provides several services, namely HTTP, FTP, TFTP, MSSQL, MySQL and SIP (VoIP).
We have installed this to expose more services after noticing failing probes on some of the associated ports of FTP, MySQL, MSSQL and SIP services. 
Enabling MSSQL on a Linux server might raise suspicions, however we felt this single exception to the rule might be overlooked or otherwise forgiven as a normal NAT configuration.
The number of connection requests to this port was so high we decided we had to see what attackers were looking for.

%Considering we are in a VM we might be limited in RAM, which Snort seems to require a lot of.
%The configured rules will have to be minimal to avoid exhausting resources.

\subsubsection{Auditing}
We reviewed various solutions to audit the command line after a successful break-in in our setup document, none of which seemed to match exactly what we needed.
The problem can be characterized as: ``From `stealthy', `maintained', and `works out of the box', choose two.''
Finally, after much research we produced a simple self-hiding kernel module that hooks \verb|execve| (new process) calls and logs the entire argument string for each call to syslog.
The module makes use of the \verb|jprobes| interface for kernel debugging, which turned out to be ideal for this task.
After it has been hidden using a `magic' command sequence, the module cannot be detected in \verb|lsmod| output, and is not detectable by \verb|rkhunter| or \verb|chkrootkit|, nor can its symbols be seen in kernel symbol lists exported by \verb|procfs|.
%TODO reference phrack

%For the HI phase, we would like to have a way to audit commandline interaction. 
%Various options exist, such as the user-space \verb|auditd| daemon, but to make auditing harder to detect this we feel it would be best to look for a kernel-space solution.
%An auditing module developed by the Honeynet project called Sebek can provide this, but is unmaintained and outdated.
%The grsec kernel patches also provide auditing from kernel space, but we were advised against replacing the stock kernel.

%We are still investigating the best way to achieve useful auditing.
%A self-written kernel module that hooks system calls, albeit limited, might suffice to get some much needed auditing information from the honeypot.

\section{Statistics}
\label{Statistics}
In this section we will discuss the statistical data that have been gathered from our logfiles. 
The logged connection attempts have been processed to group them by the IP address' originating country through the use of WHOIS directory service information.
It is worthwhile noting that a significant number of connecting hosts could not be referenced to WHOIS information.
In these cases we have attempted to resolve the missing information by making use of reverse DNS queries and grouping by associated top level domain.
Also note that these graphs display all logged connection attempts rather than successful connections, giving us a general view of which ports attract attention.

For clarity we have chosen to graph only the 10 most active originating countries, and make use of a logarithmic scale to represent the spectrum of port numbers.
The labels on the \(y\)-axis were chosen based on the most frequently accessed ports, with some selected labels removed to avoid clutter.

\includegraphics[width=1.7\textwidth,angle=-90]{tcp.png}

\includegraphics[width=1.7\textwidth,angle=-90]{udp.png}

\section{Attacks on the different services}
\label{Attacks}
During phase 2 we have monitored the honeypot closely and collected data about different attacks against the honeypot. In this section we will discuss the attacks against the services the honeypot was running.

\subsection{Kippo}

The first service we will be talking about is Kippo. During the interaction phase we've had 2610 login attempt on our Kippo client and 3871 connections. 54 of these attempts were successful.
We have made a top 5 of most tried usernames in these login attempts:

\begin{tabular}{|l|r|}
\hline
Username & Attempts \\ \hline
root & 2206 \\ \hline
admin & 39 \\ \hline
bin & 26 \\ \hline
w000t & 20 \\ \hline
oracle & 19 \\ \hline
jenny & 19 \\ \hline
\end{tabular}

It is not entirely unexpected that the username ``root'' is tried most often. 
The username ``w000t'' is a backdoor user created in the FAKE Apache 2.2.17 exploit.
Interesting to note is that ``jenny'' is the name of a fake employee listed on the our honeypot website.
Most usernames were tried multiple times, even if they made no sense to us for instance ``msnayeem\_iitd'' tried to log in four times.

These are the top 5 passwords tried:

\begin{tabular}{|l|r|}
\hline
Password & Attempts \\ \hline
  & 55 \\ \hline
123456 & 28 \\ \hline
111111 & 20 \\ \hline
root & 16 \\ \hline
admin & 16 \\ \hline
\end{tabular}

As you can see, the most tried password is a blank password. The other passwords in this top 5 are all standard password that can be found in a dictionary attack. It is nice to note that 123456 was accepted by our system.

An interesting situation was someone that tried to log in to the honeypot with the user root, just trying a list of years (1960-1980). It seems like an unusual brute force attack. Assuming someone used their birthday as a password, which still happens a lot, it is unlikely that this someone just used their birth year and not the whole birthday. However brute forcing for the entire birthday takes way to long. So for a quick and easy brute force it is unlikely to yield results.

\subsection{HTTP}
In three weeks, a total of 76002 HTTP-requests were received by our Apache server originating from 176 unique IP addresses. 4369 of these requests were 404'd.

These 404's are interesting because they show us what addresses the attackers were trying to access. For instance, we see a lot of scans for phpMyAdmin (phpMyAdmin-3.3.10.5, phpMyAdmin-3.4.7, phpMyAdmin-3.4.6 etcetera). They also searched non existing folders for these phpMyAdmin files.

Also several client IDs are visible in these logs:

\begin{itemize}
\item \emph{``-''} (3432 attempts)

This is a blank user/client ID. Which seems to be a popular client to use when you want to find weak software on a server.

\item ``MiRRORS'' (310 attempts)

Due to the well-chosen name we could not find additional information on this client. Mirrors is a frequently used word on the internet.

\item ``ZmEu'' (41 attempts)

The attackers using ``ZmEu'' found one valid path: /phpmyadmin/, however he did not exploit this. 
It is interesting to see that both several \emph{''-''} and several ``ZmEu'' users are scanning for ``/w00tw00t.at.blackhats.Romanian.anti-sec:)'' and ``/w00tw00t.at.ISC.SANS.DFind:)'' These clients are alleged to be for reconnaissance purposes.

\item ``Morfeus Fucking Scanner'' (4 attempts)

``Morfeus Fucking Scanner'' was used by one attacker, that solely searched for /user/soapCaller.bs. Multiple sources say this might be a Drupal exploit.
\end{itemize}

We expected to see several occurrences of ``/robots.txt'' because this is a common way to start when using a search engine crawler, but none were found.

\subsection{Dionaea}
As mentioned in Section \ref{Setup} Dionaea runs several services. In the table below we will show how many times those services got connected.

\begin{tabular}{|l|c|r|}
\hline
Service & Port & Number of connections \\ \hline
MSSQL  & 1433 & 1632 \\ \hline
MS SMB filesharing & 445 & 67 \\ \hline
FTP & 21 & 55 \\ \hline
MySQL & 3306 & 54 \\ \hline
Session Initiation Protocol & 5060 & 9 \\ \hline
\end{tabular}

As we looked further into the Dionaea logfiles we found that MSSQL was connected to from 95.82.11.28, which is in Iran, 1414 times. However, the SMB filesharing, FTP and SIP were mostly connected from one IP, namely 130.89.165.58. This IP is from Twente University, so probably it was another student.

\section{Connection between the attacks}
\label{Connection}
In this section we will correlate the data about attacks against one service against the data of attacks of a different service. We will try to look for a pattern or similarities in this data and determine if the attack was more sophisticated.

Three clients connected to both HTTP and Kippo, which raised our suspicion:
\begin{itemize}
\item 58.17.52.14

This citizen of the Peoples Republic of China poked every possible entrance, in two consequential days: he made a connection to Kippo and disconnected immediately. He also requested the main / folder via HTTP, using the client "-". Scanlogd also reported scans on port 80, 8080, 3128, 9415, 8909, 9592, 1080, 443. Unfortunately for us we were not attractive for him to return and actually break in.

\item 61.190.172.2

On four occasions this inhabitant of China probed us, split over 4 weeks. He also just connected to Kippo (suricata reported a SSH-scan) and disconnected: he did this for forty times in 3 minutes. Interestingly enough he was one of the users of the Morfeus Fucking Scanner client: only requesting the /user/soapCaller.bs file.

\item 125.88.75.199

The last prober was also from China. On November 22th he connected to Kippo to automaticly disconnect again. He also requested the / folder via HTTP with the "-" client. He was only seen at this occasion.

\end{itemize}

\section{Conclusion}
\label{Conclusion}
The Setup we used for this honeypot seems to be sufficient. In both phases of the research we did not lose control over our honeypot, which was our main worry. We also decided not to advertise, which gave our honeypot less traffic, but a more realistic outcome.

From the Section Attacks we can conclude that our honeypot was not as popular as we had hoped it would be. We blame this on not advertising our IP. The numbers of attacks on us were therefore not high and neither was the depth of the attacks.

Only three real scans took place, as reported in the Section Connections between attacks. Unfortuntaly we did not make our honeypot interesting enough for them to return and actually break in.

Our chose for swiching from a Low Interaction to a High Interaction honeypot seems like a good tactic, but due to downtime and bugged software, we did not have enough time.

Concluding: we did see some people checking us out, but noone actually broke in. For further research we suggest using the same setup, but having the server online for a much longer time.

\section{Discussion}
\label{Discussion}
In this section we would like to discuss some issues we encountered during the honeypot project.

\begin{itemize}
\item Low/High interaction phase

To see what the best approach is to a honeypot project, we decided to split the project in a Low and a High interaction phase. Unfortunately the High interaction phase yielded not much more results than the low interaction phase. Additionally we were a little late in switching to the HI phase. So in hindsight we don't think it was worth switching to the high interaction phase.

\item Lack of interest

We decided to make this honeypot as realistic as possible, this meant: not advertising our honeypot on IRC-channels or 4Chan. Our motivation was that in non-honeypot circumstances this would also not be done. So for example advertising the IP address of the honeypot would attract a lot more hackers than a normal server would. So the results would not be on par with attacks against a normal server. In the High interaction phase we had a website that attracted people to our honeypot, which is a more realistic approach to attracting hackers.

\item DenyHosts

To make sure our real SSH port would not be brute forced, we added DenyHost to the port. We soon realized that this was a very effective counter measurement versus intruders: both Bart and Joost were blocked due to their lack of Linux knowledge. Luckily Mark and Aram were able to unblock the other two.

\item Linux knowledge

Two of the team members had limited Linux experience. It took some time for them to understand how Linux commandline works. Fortunately we had two guru's in the group that were able to teach the newbies.

\item Google Hack Honeypot

The Google Hack Honeypot was supposed to attract hackers. Unfortunately it soon crashed due to unknown configuration errors. Because the results in the period the tool was running were non-existent and we were switching to HI later on, we decided not to use it.

\end{itemize}

%%% REFS %%%

\bibliographystyle{plain} % amsalpha
\bibliography{ref_honeypot}

\end{document}